---
title: "Lab 13 - Colonizing Mars"
author: "Cynthia Jiao"
date: "3/31/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(DT)
library(tidyverse) 
library(MASS)
library(ggplot2)
```

### Exercise 1 Simulating Our Colonists

The distribution of age resembles a normal distribution, because rnorm function is to extract data from a normal distribution. 
```{r}

set.seed(123)
age <- rnorm(100, mean = 30, sd = 5)

df_colonists <- data.frame(id = 1:100, # this is our id column
                            age = age ) # add the age column here


ggplot(df_colonists, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "white") +
  labs(title = "Age Distribution of Colonists",
       x = "Age",
       y = "Count") +
  theme_minimal()


```


```{r}


set.seed(1)
roleA <- rep(c("engineer", "scientist", "medic"), 
             length.out = 100)
# this works if we want to set a specific number of each role
roleB <- rep(c("engineer", "scientist", "medic"), 
             each = 34, 
             length.out = 100)

roleC <- rep(c("engineer", "scientist", "medic"), 
             times = c(33, 33, 33), 
             length.out = 100)
             
# if you want to use sampling weights
roleD <- sample(c("engineer", "scientist", "medic"), 
                replace = TRUE, 
                size = 100, prob = c(1,1,1))

# if you want to randomly shuffle

roleE <- sample(roleB, size = 100, replace = FALSE)

## Hey Mason, I am not understanding how the codes above work, so I just did the below, which seems to achieve the same goal we intended?

## adding randomized role variable to df_colonist
roles <- c("Scientist", "Farmer", "Medic", "Engineer", "Cook")

# Randomly assign one role to each colonist
set.seed(123)  # for reproducibility
df_colonists$role <- sample(roles, size = nrow(df_colonists), replace = TRUE)


## in addition to engineer, scientist, and medic, a colony probably needs workers to get the work done, so adding workers to the role variable...

set.seed(42)

# create roles (25 of each for 100 people)
roles <- rep(c("Engineer", "Worker", "Medic", "Scientist"), each = 25)

# randomly shuffle
df_colonists$role <- sample(roles)


## create the MARSGAR variable

set.seed(123)

df_colonists$marsgar <- runif(100, min = 0, max = 100)


```


### Exercise 2

#creating variables that have specific correlations

```{r}

# Simulate technical skills based on age

# Set the seed for reproducibility
set.seed(123)

# Create a variable for technical skills
df_colonists$technical_skills <- 2 * df_colonists$age + rnorm(100, mean = 0, sd = 1)

# visualize the relationship

ggplot(df_colonists, aes(x = age, y = technical_skills)) +
  geom_jitter(color = "purple", size = 1.5) +
  geom_smooth(method = "lm", color = "black") +
  labs(title = "Scatter Plot of Colonist Age and technical skills",
       x = "Age",
       y = "Skills") +
  theme_minimal()

```
Engineer probably have the highest problem solving abilities, since solving problem is really a large part of their job... Then scientist and medic have second and third highest problem solving abilities, followed by workers, who need more instructions to get work done.
```{r}
##since I have 4 levels in role...

set.seed(123)
df_colonists$problem_solving[df_colonists$role == "Engineer"]  <- rnorm(sum(df_colonists$role == "Engineer"),  mean = 100, sd = 10)
df_colonists$problem_solving[df_colonists$role == "Scientist"] <- rnorm(sum(df_colonists$role == "Scientist"), mean = 80,  sd = 10)
df_colonists$problem_solving[df_colonists$role == "Medic"]     <- rnorm(sum(df_colonists$role == "Medic"),     mean = 70,  sd = 10)
df_colonists$problem_solving[df_colonists$role == "Worker"]    <- rnorm(sum(df_colonists$role == "Worker"),    mean = 60,  sd = 10)


```

### Exercise 3

#stimulate data using multivariate rnorm => mvrnorm (MASS)

```{r}

?mvrnorm

mean_traits <- c(50, 50)
cov_matrix <- matrix(c(100, 50, 
                       50, 100), ncol = 2)

# Generate correlated data

traits_data <- mvrnorm(n = 100, mu = mean_traits, Sigma = cov_matrix, empirical = FALSE)

colnames(traits_data) <- c("resilience", "agreeableness")

df_colonists <- cbind(df_colonists, traits_data)

```

# stimulating Big Five

```{r}

  cor_matrix_bigfive <- matrix(c(
  1.0000, 0.2599, 0.1972, 0.1860, 0.2949,
  0.2599, 1.0000, 0.1576, 0.2306, 0.0720,
  0.1972, 0.1576, 1.0000, 0.2866, 0.1951,
  0.1860, 0.2306, 0.2866, 1.0000, 0.1574,
  0.2949, 0.0720, 0.1951, 0.1574, 1.0000
  ), nrow=5, ncol=5, byrow=TRUE,
  dimnames=list(c("EX", "ES", "AG", "CO", "OP"), 
        c("EX", "ES", "AG", "CO", "OP")))

# Generate correlated data
mu_bigfive <- rep(0, 5)

big_five_data <- mvrnorm(n = 100, mu = mu_bigfive, Sigma = cor_matrix_bigfive, empirical = FALSE)

colnames(big_five_data) <- c("openness", "agreeableness", "conscientiousness", "extraversion", "neuroticism") ## this agreeableness here will be agreeableness.1, since we had another agreeableness before

df_colonists <- cbind(df_colonists, big_five_data)


```

### Excercise 4

#replicate the colonies

```{r}

set.seed(12)

# simulate 100 planets
colonies_list <- replicate(
  n = 100,
  expr = {
    scores <- mvrnorm(n = 100, mu = mu_bigfive, Sigma = cor_matrix_bigfive)
    df <- as.data.frame(scores)
    colnames(df) <- c("Extraversion", "Neurotism", 
                      "Agreeableness", "Conscientiousness", "Openness")
    return(df)
  },
  simplify = FALSE
)

df_all <- bind_rows(
  lapply(1:100, function(i) {
    colonies_list[[i]] %>%
      mutate(planet = i, colonist_id = 1:100)
  })
)

summary_df <- df_all %>%
  group_by(planet) %>%
  summarise(
    mean_extraversion = mean(Extraversion),
    sd_extraversion = sd(Extraversion),
    correlation_with_openness = cor(Extraversion, Openness)
  )

print(summary_df)
```

The distribution of the r between extraversion and openness is approaching a normal distribution, because we stimulate the data with a given correlation of 0.2949, and the more sample we randomly draw from the population, the more this distribution is going to resemble a normal distribution. 
```{r}

ggplot(summary_df, aes(x = correlation_with_openness)) +
  geom_histogram(binwidth = 0.02, fill = "skyblue", color = "white") +
  labs(
    title = "Distribution of Correlations Between Extraversion and Openness",
    x = "Correlation Coefficient (r)",
    y = "Number of Colonies"
  ) +
  theme_minimal()


```

